---
title: "Calculate Income-to-Needs at T1"
author: "Lucy King"
output: 
  github_document:
    toc: true
    toc_depth: 6
---


```{r}
##Libraries
library(tidyverse)

##Parameters
#files
income_t1_file <- "~/Box/lucy_king_files/ELS/income_TBM/data/final_data/income_t1.csv"

included_ids_file <- "~/Box/lucy_king_files/ELS/income_TBM/data/final_data/included_ids.csv"
```

## calculate income-to-needs: Santa Clara County Low Income Thresholds 
```{r}
#income reported in bins, used median for income to needs as in Yu et al. (2017) developmental science
inr <-
  read_csv(income_t1_file) %>% 
  select(
    ELS_ID,
    household_income = Household_Income,
    ppl_in_home = pplinhome_1_TEXT
  ) %>% 
  filter(
    !is.na(household_income)
    ) %>% 
  mutate(
    household_income = as.character(household_income),
    household_income_numeric = as.numeric(
      recode(
        household_income,
        "1" = "5000",
        "2" = "7500",
        "3" = "12500",
        "4" = "20000",
        "5" = "30000",
        "6" = "42500",
        "7" = "62500",
        "8" = "87500",
        "9" = "125000",
        "10" = "150000"
      )
    ),
#santa clara country low-income limit : 80% of median income (ratios <= 1 are therefore "low income")
#https://www.huduser.gov/portal/datasets/il/il2017/2017summary.odn
    SC_lowincome_limit = as.numeric(
      recode(
        ppl_in_home,
        "1" = "59350",
        "2" = "67800",
        "3" = "76300", 
        "4" = "84750",
        "5" = "91500",
        "6" = "98350",
        "7" = "105100",
        "8" = "111900",
        .default = "111900"
      )
    ),
    income_needs = household_income_numeric / SC_lowincome_limit
  )
```

## calculate income-to-needs: Santa Clara County Very Low Income Thresholds 
```{r}
inr <-
  inr %>% 
  mutate(
    SC_verylowincome_limit = as.numeric(
      recode(
        ppl_in_home,
        "1" = "41800",
        "2" = "47800",
        "3" = "53750",
        "4" = "59700",
        "5" = "64500",
        "6" = "69300",
        "7" = "74050",
        "8" = "78850",
        .default = "78850"
      )
    ),
    income_needs_SC_vlow = household_income_numeric / SC_verylowincome_limit
  )
```

## calculate income-to-needs: Santa Clara County Extremely Low Income Thresholds 
```{r}
inr <-
  inr %>% 
  mutate(
    SC_extrmlowincome_limit = as.numeric(
      recode(
        ppl_in_home,
        "1" = "25100",
        "2" = "28650",
        "3" = "32250",
        "4" = "35800",
        "5" = "38700",
        "6" = "41550",
        "7" = "44400",
        "8" = "47300",
        .default = "47300"
      )
    ),
    income_needs_SC_extlow = 
      household_income_numeric / SC_extrmlowincome_limit
  )
```

## calculate income-to-needs: Federal Low Income Thresholds 
```{r}
inr <-
  inr %>% 
  mutate(
    federal_lowincome_limit = as.numeric(
      recode(
        ppl_in_home,
        "1" = "12488",
        "2" = "15877",
        "3" = "19515",
        "4" = "25094",
        "5" = "29714",
        "6" = "33618",
        "7" = "38173",
        "8" = "42684",
        "9" = "50681",
        .default = "50681"
      )
    )
  ) %>% 
  mutate(
    income_needs_fed_poverty = 
      household_income_numeric / federal_lowincome_limit
  )
    
```

## subset to only participants included in paper
```{r}
included_ids <-
  read_csv(included_ids_file) %>% 
  pull(ELS_ID)

inr <-
  inr %>% 
  filter(ELS_ID %in% included_ids)
```

## visualize
```{r}
inr %>% 
  select(
    income_needs,
    income_needs_SC_vlow,
    income_needs_SC_extlow,
    income_needs_fed_poverty
  ) %>% 
  gather(key, val) %>% 
  mutate(
    key = factor(
      key,
      levels = c(
        "income_needs",
        "income_needs_SC_vlow",
        "income_needs_SC_extlow",
        "income_needs_fed_poverty"
      ),
      labels = c(
        "Santa Clara County - Low Income Threshold",
        "Santa Clara County - Very Low Income Threshold",
        "Santa Clara County - Extremely Low Income Threshold",
        "US Census National - Poverty Threshold"
      )
    )
  ) %>% 
  ggplot(aes(val)) +
  geom_histogram() +
  geom_vline(xintercept = 1, color = "red") +
  scale_x_continuous(breaks = seq.int(0, 8, 1)) +
  facet_wrap(.~key, nrow = 4, ncol = 1, scales = "free") +
  theme_minimal() +
  theme(
    legend.title = element_blank(),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    legend.text = element_text(size = 18),
    strip.text = element_text(size = 18),
    legend.key.size = unit(2, 'lines')
  ) +
  labs(
    x = "Income-to-needs ratio"
  )

ggsave(
  "~/Box/lucy_king_files/ELS/income_TBM/plots/inr_histograms.png",
  height = 9,
  width = 7
)
```

## counts: SC low income, SC very low income, SC extremely low income, federal poverty
```{r}
inr %>% 
  count(income_needs < 1) %>% 
  mutate(per = n / sum(n))

inr %>% 
  count(income_needs_SC_vlow < 1) %>% 
  mutate(per = n / sum(n))

inr %>% 
  count(income_needs_SC_extlow < 1) %>% 
  mutate(per = n / sum(n))

inr %>% 
  count(income_needs_fed_poverty < 1) %>% 
  mutate(per = n / sum(n))

```

```{r}
inr %>% 
  arrange(desc(income_needs)) %>% 
  select(income_needs, income_needs_fed_poverty)
```

## Children in poverty in Bay Area communities
https://www.census.gov/quickfacts/fact/map/santaclaracountycalifornia/IPE120218

Bay Area = nine counties: Alameda, Contra Costa, Marin, Napa, San Mateo, Santa Clara, Solano, Sonoma, and San Francisco. 

```{r}
bay_area <- 
  tibble(
  county = c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma",
    "San Francisco"
  ),
  fed_poverty_rate = c(
    .09,
    .078,
    .066,
    .088,
    .068,
    .073,
    .079,
    .099,
    .101
  ),
  population = c(
    1666753,
    1150215,
    259666,
    139417,
    769545,
    1937570,
    446610,
    499942,
    883305
  ),
  poverty_individuals = fed_poverty_rate * population
)

bay_area %>% 
  summarise(
    total_poverty = sum(poverty_individuals),
    total_population = sum(population),
    overall_fed_rate = total_poverty / total_population
  )
```


