---
title: "Effects of INR on T1 symptoms"
author: "Lucy King"
output: 
  github_document:
    toc: true
    toc_depth: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
##Libraries
library(tidyverse)

##Parameters

#files____________________________________________________________________

#parent-reported symptoms
t1_cbcl_file <- "~/Desktop/ELS/income_TBM/data/osf/ELS_T1S1_CBCL_old_1991_Final Profile Data_ID 1-218.sav" 

#child-reported symptoms
t1_ysr_file <- "~/Desktop/ELS/income_TBM/data/osf/ELS_T1S1_YSR_old_1991_Final Profile Data_ID 1-218.sav" 

#extracted estimates for all 46 ROIs identified in sex*income model
roi_inc_int_file <- "~/Desktop/ELS/income_TBM/data/osf/inc_interac_46rois_all.xlsx" 

#income-to-needs ratios
inr_file <- "~/Desktop/ELS/income_TBM/data/osf/income_to_needs.csv"

#covariates used in income TBM model
tbm_covariates_file <- "~/Desktop/ELS/income_TBM/data/osf/tbm_cov_income.csv"
```

##Join data
```{r}
roi_symptoms <-
  readxl::read_xlsx(roi_inc_int_file, sheet = "Vols") %>% 
  left_join(read_csv(tbm_covariates_file), by = "ELS_ID") %>% 
  left_join(read_csv(inr_file), by = "ELS_ID") %>% 
  left_join(
    haven::read_sav(t1_cbcl_file, user_na = 999) %>% 
      mutate(
        ELS_ID = as.integer(ELS_ID),
        cbcl_externalizing = as.integer(
          T1_CBCL_old_Externalizing_Problems_Total_Score
        ),
        cbcl_externalizing = if_else(
          cbcl_externalizing == 999, NA_integer_, cbcl_externalizing
        )
      ) %>% 
      select(
        ELS_ID,
        cbcl_externalizing
      ), 
    by = "ELS_ID"
  ) %>% 
  left_join(
    haven::read_sav(t1_ysr_file, user_na = 999) %>% 
      mutate(
        ELS_ID = as.integer(ELS_ID),
        ysr_internalizing = as.integer(
          T1_YSR_old_Internalizing_Problems_Total_Score
        ),
        ysr_internalizing = if_else(
          ysr_internalizing == 999, NA_integer_, ysr_internalizing
        )
      ) %>% 
      select(
        ELS_ID,
        ysr_internalizing
      )
  )
```

##Create coding for sex and race
```{r}
roi_symptoms <-
  roi_symptoms %>% 
  mutate(
    male_simple_effect = recode(
      Male,
      "1" = "0",
      "0" = "1"
    ),
    female_simple_effect = recode(
      Male,
      "1" = "1",
      "0" = "0"
    ),
    sex_effect_coded = recode(
      Male,
      "1" = "1", #male
      "0" = "-1" #female
    ),
    white_effect_coded = recode(
      White,
      "1" = "1",
      "0" = "-1"
    )
  ) 
```

##Create separate tibbles for boys and girls
```{r}
#male only tibble
roi_symptoms_male <- 
  roi_symptoms %>% 
  filter(Male == 1)

#female only tibble
roi_symptoms_female <- 
  roi_symptoms %>% 
  filter(Male == 0)
```

##Distributions of symptoms
```{r symptom_hist}
#externalizing
roi_symptoms %>% 
  ggplot(aes(x = cbcl_externalizing)) +
  geom_histogram(bins = 30, position = "dodge", fill = "gray", color = "black") +
  theme_bw()

roi_symptoms_male %>% 
  ggplot(aes(x = cbcl_externalizing)) +
  geom_histogram(bins = 30, position = "dodge", fill = "gray", color = "black") +
  theme_bw()

roi_symptoms_female %>% 
  ggplot(aes(x = cbcl_externalizing)) +
  geom_histogram(bins = 30, position = "dodge", fill = "gray", color = "black") +
  theme_bw()

# internalizing
roi_symptoms %>% 
  ggplot(aes(x = ysr_internalizing)) +
  geom_histogram(bins = 30, position = "dodge", fill = "gray", color = "black") +
  theme_bw()

roi_symptoms_male %>% 
  ggplot(aes(x = ysr_internalizing)) +
  geom_histogram(bins = 30, position = "dodge", fill = "gray", color = "black") +
  theme_bw()

roi_symptoms_female %>% 
  ggplot(aes(x = ysr_internalizing)) +
  geom_histogram(bins = 30, position = "dodge", fill = "gray", color = "black") +
  theme_bw()
```

##Externalizing
```{r}
#main effect
mod_extern_income = lm(cbcl_externalizing ~ sqrt(income_needs), data = roi_symptoms)

summary(mod_extern_income)

AIC(mod_extern_income)

confint(mod_extern_income)

#interaction with sex 
mod_sex_extern_income = 
  lm(cbcl_externalizing ~ sqrt(income_needs) * sex_effect_coded, data = roi_symptoms)

summary(mod_sex_extern_income)

AIC(mod_sex_extern_income)

confint(mod_sex_extern_income)

#interaction with TBM covariates
mod_sex_extern_income_cov = 
  lm(cbcl_externalizing ~ sqrt(income_needs) * sex_effect_coded + white_effect_coded + Tanner + Age, data = roi_symptoms)

summary(mod_sex_extern_income_cov)

confint(mod_sex_extern_income_cov)

#simple effects

#males
mod_male_extern_income = 
  lm(cbcl_externalizing ~ sqrt(income_needs) * male_simple_effect, data = roi_symptoms)

summary(mod_male_extern_income)

confint(mod_male_extern_income)

mod_male_extern_income_cov = 
  lm(cbcl_externalizing ~ sqrt(income_needs) * male_simple_effect + white_effect_coded + Tanner + Age, data = roi_symptoms)

summary(mod_male_extern_income_cov)

confint(mod_male_extern_income_cov)

summary(lm(cbcl_externalizing ~ sqrt(income_needs), data = roi_symptoms_male))

summary(lm(cbcl_externalizing ~ sqrt(income_needs) + white_effect_coded + Tanner + Age, data = roi_symptoms_male))


#females
mod_female_extern_income = 
  lm(cbcl_externalizing ~ sqrt(income_needs) * female_simple_effect, data = roi_symptoms)

summary(mod_female_extern_income)

confint(mod_female_extern_income)

mod_female_extern_income_cov = 
  lm(cbcl_externalizing ~ sqrt(income_needs) * female_simple_effect +  white_effect_coded + Tanner + Age, data = roi_symptoms)

summary(mod_female_extern_income_cov)

confint(mod_female_extern_income_cov)

summary(lm(cbcl_externalizing ~ sqrt(income_needs), data = roi_symptoms_female))

summary(lm(cbcl_externalizing ~ sqrt(income_needs) +  white_effect_coded + Tanner + Age, data = roi_symptoms_female))
```

##Internalizing
```{r}
#main effect
mod_intern_income = lm(ysr_internalizing ~ sqrt(income_needs), data = roi_symptoms)

summary(mod_intern_income)

AIC(mod_intern_income)

confint(mod_intern_income)

#interaction with sex 
mod_sex_intern_income = 
  lm(ysr_internalizing ~ sqrt(income_needs) * sex_effect_coded, data = roi_symptoms)

summary(mod_sex_intern_income)

AIC(mod_sex_intern_income)

confint(mod_sex_intern_income)

#interaction with TBM covariates
mod_sex_intern_income_cov = 
  lm(ysr_internalizing ~ sqrt(income_needs) * sex_effect_coded + white_effect_coded + Tanner + Age, data = roi_symptoms)

summary(mod_sex_intern_income_cov)

confint(mod_sex_intern_income_cov)
```

##Visualize externalizing sex differences
```{r}
#externalizing 
roi_symptoms %>%
  mutate(
    Sex = factor(
      Male,
      levels = c("0", "1"),
      labels = c("Female", "Male")
    )
  ) %>% 
  ggplot(aes(income_needs, cbcl_externalizing, color = fct_rev(Sex))) +
  geom_point(size = 2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ sqrt(x), size = 2, se = FALSE) +
  scale_y_continuous(breaks = seq(0, 30, 5)) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 24),
    axis.title = element_text(size = 28),
    axis.text = element_text(size = 24)
  ) +
  labs(
    color = NULL,
    x = "Income-to-needs ratio",
    y = "Externalizing symptoms"
  )

ggsave(
  "~/Desktop/ELS/income_TBM/income_TBM_sync/plots/inr_sex_externalizing.jpeg",
  width = 8,
  height = 8
  )
```

```{r}
#internalizing 
roi_symptoms %>%
  mutate(
    Sex = factor(
      Male,
      levels = c("0", "1"),
      labels = c("Female", "Male")
    )
  ) %>% 
  ggplot(aes(income_needs, ysr_internalizing, color = fct_rev(Sex))) +
  geom_point(size = 2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ sqrt(x), size = 2, se = FALSE) +
  scale_y_continuous(breaks = seq(0, 45, 5)) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 24),
    axis.title = element_text(size = 28),
    axis.text = element_text(size = 24)
  ) +
  labs(
    color = NULL,
    x = "Income-to-needs ratio",
    y = "Internalizing symptoms"
  )

ggsave(
  "~/Desktop/ELS/income_TBM/income_TBM_sync/plots/inr_sex_internalizing.jpeg",
  width = 8,
  height = 8
  )
```

##Visualize mediation
```{r}
roi_symptoms %>% 
  filter(Male == 1) %>% 
  ggplot(aes(income_needs, Neg_18)) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  geom_point(size = 4, color = "blue", alpha = 1/2) +
  geom_smooth(
    method = "lm", 
    formula = y ~ sqrt(x), 
    se = FALSE, 
    size = 3,
    color = "blue"
    ) +
  theme_minimal() +
  theme(
    legend.text = element_text(size = 24),
    axis.title = element_text(size = 28),
    axis.text = element_text(size = 24)
  ) +
  labs(
    x = "Income-to-needs ratio",
    y = "Left MTG volume"
  )

ggsave(
  "~/Desktop/ELS/income_TBM/income_TBM_sync/plots/inr_mtg_boys.jpeg",
  width = 8,
  height = 8
  )
```

```{r}
roi_symptoms %>% 
  filter(Male == 1) %>% 
  ggplot(aes(Neg_18, cbcl_externalizing)) +
  geom_point(size = 4, color = "blue", alpha = 1/2) +
  geom_smooth(
    method = "lm", 
    se = FALSE, 
    size = 3,
    color = "blue"
    ) +
  theme_minimal() +
  theme(
    legend.text = element_text(size = 24),
    axis.title = element_text(size = 28),
    axis.text = element_text(size = 24)
  ) +
  labs(
    x = "Left MTG volume",
    y = "Externalizing symptpms"
  )


ggsave(
  "~/Desktop/ELS/income_TBM/income_TBM_sync/plots/mtg_extern_boys.jpeg",
  width = 8,
  height = 8
  )
```

##Visualize hippocampus sex differences
```{r}
roi_symptoms %>%
  mutate(
    Sex = factor(
      Male,
      levels = c("0", "1"),
      labels = c("Female", "Male")
    )
  ) %>% 
  ggplot(aes(income_needs, Neg_39, color = fct_rev(Sex))) +
  geom_point(size = 2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ sqrt(x), size = 2, se = FALSE) +
  scale_color_manual(values = c("blue", "red")) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 24),
    axis.title = element_text(size = 28),
    axis.text = element_text(size = 24)
  ) +
  labs(
    color = NULL,
    x = "Income-to-needs ratio",
    y = "Left MTG/hippocampus/\nCGH volume"
  )

ggsave(
  "~/Desktop/ELS/income_TBM/income_TBM_sync/plots/inr_sex_left_hipp.jpeg",
  width = 8,
  height = 8
  )
```

##Sensitivity analyses

###Compare sample to those missing brain data on symptoms
```{r}
symptoms_roi <-
  haven::read_sav(t1_cbcl_file, user_na = 999) %>% 
  mutate(
    ELS_ID = as.integer(ELS_ID),
    cbcl_externalizing = as.integer(
      T1_CBCL_old_Externalizing_Problems_Total_Score
    ),
    cbcl_externalizing = if_else(
      cbcl_externalizing == 999, NA_integer_, cbcl_externalizing
    )
  ) %>% 
  select(ELS_ID, cbcl_externalizing) %>% 
  left_join(
    haven::read_sav(t1_ysr_file, user_na = 999) %>% 
      mutate(
        ELS_ID = as.integer(ELS_ID),
        ysr_internalizing = as.integer(
          T1_YSR_old_Internalizing_Problems_Total_Score
        ),
        ysr_internalizing = if_else(
          ysr_internalizing == 999, NA_integer_, ysr_internalizing
        )
      ) %>% 
      select(ELS_ID, ysr_internalizing),
    by = "ELS_ID"
  ) %>% 
  left_join(
    read_csv(tbm_covariates_file) %>% 
      select(
        ELS_ID,
        included = ICV
      ), 
    by = "ELS_ID"
  ) %>% 
  mutate(
    included = if_else(
      !is.na(included), 
      "included", "not"
    )
  )

t.test(symptoms_roi$ysr_internalizing ~ symptoms_roi$included)
t.test(symptoms_roi$cbcl_externalizing ~ symptoms_roi$included)
```

###linear effects of INR on externalizing symptoms 
```{r}
#linear main effect
mod_extern_income_lin = lm(cbcl_externalizing ~ income_needs, data = roi_symptoms)

summary(mod_extern_income_lin)

AIC(mod_extern_income_lin)

confint(mod_extern_income_lin)

#linear interaction with sex
mod_sex_extern_income_lin = 
  lm(cbcl_externalizing ~ income_needs * sex_effect_coded, data = roi_symptoms)

summary(mod_sex_extern_income_lin)

AIC(mod_sex_extern_income_lin)

confint(mod_sex_extern_income_lin)

#simple effects

#males
mod_male_extern_income_lin = 
  lm(cbcl_externalizing ~ income_needs * male_simple_effect, data = roi_symptoms)
summary(mod_male_extern_income_lin)
confint(mod_male_extern_income_lin)

summary(lm(cbcl_externalizing ~ income_needs, data = roi_symptoms_male))

#females
mod_female_extern_income_lin = 
  lm(cbcl_externalizing ~ income_needs * female_simple_effect, data = roi_symptoms)
summary(mod_female_extern_income_lin)
confint(mod_female_extern_income_lin)

summary(lm(cbcl_externalizing ~ income_needs, data = roi_symptoms_female))
```

###linear effects of INR on internalizing symptoms 
```{r}
#linear main effect
mod_intern_income_lin = lm(ysr_internalizing ~ income_needs, data = roi_symptoms)

summary(mod_intern_income_lin)

AIC(mod_intern_income_lin)

confint(mod_intern_income_lin)

#linear interaction with sex
mod_sex_intern_income_lin = 
  lm(ysr_internalizing ~ income_needs * sex_effect_coded, data = roi_symptoms)

summary(mod_sex_intern_income_lin)

AIC(mod_sex_intern_income_lin)

confint(mod_sex_intern_income_lin)

#simple effects

#males
mod_male_extern_income_lin = 
  lm(cbcl_externalizing ~ income_needs * male_simple_effect, data = roi_symptoms)
summary(mod_male_extern_income_lin)
confint(mod_male_extern_income_lin)

summary(lm(cbcl_externalizing ~ income_needs, data = roi_symptoms_male))

#females
mod_female_extern_income_lin = 
  lm(cbcl_externalizing ~ income_needs * female_simple_effect, data = roi_symptoms)
summary(mod_female_extern_income_lin)
confint(mod_female_extern_income_lin)

summary(lm(cbcl_externalizing ~ income_needs, data = roi_symptoms_female))
```